Out-of-Sample Comparisons of Overfit Models

### 2013-12-16 paper.tex
Remove end-of-line whitespace

Both models are estimated using \ols\ with a fixed window:
\[
\hat y_{i,t+\h} = x_{it}'\hat{\theta}_{it}, \qquad i=1,2, \qquad t
= R+1,\dots,T-h
\]
with
\begin{equation*}
  \bh{it} =
  \begin{cases}
    \big(\sum_{s=1}^{R-\h} x_{is}x_{is}'\big)^{-1} \sum_{s=1}^{R-\h}
    x_{is} y_{s+\h} & t \leq T \\
    \big(\sum_{s=1}^{T-\h} x_{is}x_{is}'\big)^{-1}
    \sum_{s=1}^{T-\h} x_{is} y_{s+\h} & t > T.
  \end{cases}
\end{equation*}
The observed \oos\ average is denoted $\oosA$: \[\oosA \equiv P^{-1}
\oosSum{t}{1} D_t\]  with
\[
D_t \equiv L(y_{t+\h} - x_{1t}'\bh{1t}) - L(y_{t+\h} - x_{2t}'\bh{2t}).
\]
The \oos\ statistic of interest is $\oost \equiv \sqrt{P} \oosA /
\sh$, for $\sh^2$ some (possibly Heteroskedasticity- and
Autocorrelation-Consistent, or \hac) estimator of the asymptotic
variance of $\oosA$.

As discussed above, the forecaster's goal is to choose the model with
lower average future loss; i.e. to determine whether $\oosB$ is
positive or negative,with
\[
\oosB \equiv Q^{-1} \sum_{t=T+1}^{T+Q} D_t.
\]
Note that the coefficients are reestimated with the full dataset.
Since the forecaster's information set, $\mathcal{F}_T$, does not
contain $\oosB$, he or she should decide between the models based on
the conditional expectation, $\E_T \oosB$. If this quantity is
positive, the first model is expected to perform the worst in the
future, and if it is negative the second is.

@@

\begin{asmp}\label{asmp-3}
  The loss function $L$ is continuous and has finite left and right
  derivatives almost everywhere.  Also, $L(0) = 0$, and there is a
  constant $B_L$ such that $\|D_t\|_\rho \leq B_L$ for all $t$ ($\rho$ is defined
  in Assumption~\ref{asmp-1}); $\| x_{it}' \hat{\theta}_{it} \|_2 \leq
  B_L$ and $\| x_{it}' \hat{\theta}_{it} \|_2 \leq B_L$; and
  \begin{equation*}
    \| L'(y_{t+h} - x_{it}'\tilde{\theta}_{it}) \|_2 \leq B_L
  \end{equation*}
  for $\tilde{\theta}_{it}$ a weighted average of any two estimates
  $\hat{\theta}_{it}$ and $\hat{\theta}_{it'}$.
\end{asmp}

@@

\begin{thm}\label{res-confidence-intervals}
  Suppose that Assumptions \ref{asmp-1}--\ref{asmp-4} hold, that
  $\sh^2$ is an estimator satisfying
  \[
    \sh^2 - \sigma^2 \to^p 0 \quad \text{with}\quad
    \sigma^2 \equiv \var_R(\sqrt{P} \oosA) \equiv P \E_R(\bar{D}_R -
    \E_R \bar{D}_R)^2,
  \]
  and that $\sigma^2$ is uniformly a.s. positive.  Then
  \[
  \frac{\oosA - \E_T\oosB}{\sh / \sqrt{P}}
  \to^d N(0,1).
  \]
  Consequently, each of the usual Gaussian confidence intervals,
  \begin{gather}
  [\oosA - z_{\alpha/2} \, \sh /
      \sqrt{P}, \oosA + z_{\alpha/2} \sh / \sqrt{P}],\label{interval-twosided} \\
  [\oosA - z_{\alpha} \, \sh / \sqrt{P}, +\infty), \label{interval-greater}
  \intertext{and}
(-\infty, \oosA + z_{\alpha}
      \, \sh / \sqrt{P}],
    \end{gather}
    contains $\E_T\oosB$ with probability $\alpha$ in the limit,
  with $z_{\alpha}$ the $1-\alpha$ quantile of the standard normal
  distribution.
\end{thm}
The requirement that $\sigma^2$ be uniformly a.s. positive is not
restrictive under our asymptotic theory.  Since the models'
coefficients are estimated with uncertainty in the limit, the two
models give different forecasts even if they both nest the \dgp.  This
is intuitively similar to \citepos{GiW:06} rolling-window result, but
comes from different asymptotic theory; Giacomini and White keep the
variance of the \oos\ average positive by letting $P \to \infty$ with
$R$ fixed; in this paper, $R \to \infty$ but the variance remains
positive since $K \to \infty$ (quickly) as well.

This result allows forecasters to use the \dmw\ test to conduct
inference about the models' accuracy.  Since the \oos\ average,
$\oosA$, is asymptotically normal and centered on the expected future
performance, $t$-tests constructed from the \oos\ errors necessarily
test hypotheses about $\E_T \oosB$ and not the performance of the
pseudo-true models or the unconditional performance of the models.
This application to testing can be expressed as its own theorem:
\begin{thm}\label{res:oostest}
Suppose that the conditions of Theorem \ref{res-confidence-intervals}
hold and that $\lim \Pr[\E_T \oosB \leq 0] > 0$.  Then
\begin{equation}
  \lim \Pr[P^{1/2}\oosA/\sh > z_{\alpha} \mid \E_T
  \oosB \leq 0] \leq \alpha.
\end{equation}
\end{thm}

Two intermediate Lemmas will help us understand Theorem
\ref{res-confidence-intervals} better.  Its proof consists of two
steps: we can write
\begin{equation}
  \sqrt{P} (\oosA - \E_T \oosB) = \sqrt{P} (\oosA - \E_R
  \oosA) + \sqrt{P} (\E_R \oosA - \E_T \oosB),
\end{equation}
so Theorem~\ref{res-confidence-intervals} holds if
\begin{equation}\label{coefpart}
  \E_R \oosA = \E_T \oosB + o_p(P^{-1/2}).
\end{equation}
and
\begin{equation}\label{normalpart}
  \sqrt{P}\frac{\oosA - \E_R \oosA}{\sigma} \to^d N(0,1)
\end{equation}
The first Lemma establishes~\eqref{coefpart} and the second shows that
the process $\{D_t - \E_R \oosA\}$ is an $L_2$-mixingale of size
$-1/2$ that satisfies a \clt, ensuring~\eqref{normalpart}.

@@

Note that $\E_T \oosB - \E_R \oosA$ is noise introduced by
approximating the performance of the full-sample estimates with that
of the partial-sample estimates.  Moreover, the $O_p(\sqrt{P/T})$ term
is binding; it is straightforward to find examples for which
$\sqrt{T/P}(\E_T \oosB - \E_R \oosA)$ does not converge to
zero.\footnote{Notice that, if the observations are i.i.d. and
  $(y_{t+h}, x_{it}) \sim N(0,I)$, $(\hat{\theta}_{iR} -
  \hat{\theta}_{iT}) \sim N(0, (X_{iR}'X_{iR})^{-1} -
  (X_{iT}'X_{iT})^{-1})$ given $X_{iT}$, and so
  $\var(x_{i,T+1}'\hat{\theta}_{iR} - x_{i,T+1}'
  \hat{\theta}_{iT}) \gg P K_i / (R - K_2)(T - K_2)$.}
Consequently, there is little scope to relax our assumption that
$P^2/T \to 0$.  If $\lim P^2/T$ is positive and finite then there is
the possibility of adjusting the asymptotic distribution of the test
statistic to account for this noise term as in \citet{Wes:96}, but we
leave that to future research.  If $P^2/T \to \infty$, which is
closest to common practice in applied research, the noise term
dominates the asymptotic distribution of the \oos\ $t$-test and the
test has only trivial power in the limit if it has correct size.

\begin{lem}\label{res-mixingale}
  Part 1: If Assumptions~\ref{asmp-1}--\ref{asmp-3} hold then the
  inequality
  \begin{equation}\label{mixingaleR}
    \lVert \E_{R+j-l}(D_{R+j} - \E_R D_{R+j}) \rVert_2 \leq
    \couplingConstant \; \zeta_l
  \end{equation}
  with $\zeta_l = O(l^{-1/2 - \delta})$ holds for some positive
  $\delta$, any positive $j$, and any $l$ between 0 and $j$.  As a
  consequence, $\{D_t - \E_R D_t, \mathcal F_t\}$ is an
  $L_2$-mixingale of size $-1/2$.

  \noindent Part 2: Moreover,~\eqref{normalpart} holds under
  Assumptions~\ref{asmp-1}--\ref{asmp-4} if $\sigma^2$ is
  a.s. uniformly positive.
\end{lem}

@@

Finally, for Theorem \ref{res-confidence-intervals} to be useful, we
need a consistent estimator of $\sigma^2$. Lemma
\ref{res-variance-estimator} establishes that the usual \oos\ \hac\
variance estimators are consistent.\footnote{As with
      Lemma~\ref{res-mixingale}, the proof of this lemma takes
      \citepos{JoD:00} existing consistency results
      for \ned\ processes and shows that the same arguments can be
      applied to our \oos\ mixingale.}

\begin{lem}
  \label{res-variance-estimator} Suppose Assumptions
  \ref{asmp-1}--\ref{asmp-5} hold and that $\gamma \to \infty$ and
  $\gamma/P \to 0$ as $T \to \infty$, and define
  \begin{equation*}
    \sh^2 \equiv P^{-1} \oosSum{s,t}{1} (D_t - \oosA)(D_{Rs} - \oosA)
    \vWeight.
  \end{equation*}
  Then $\sh^2 - \sigma^2 \to^p 0$.
\end{lem}

@@

The first two conditional moments of $D_t$ for $t > R$ (given
$\mathcal{F}_R$) are straightforward to derive: $\E_R D_t = -
\hat{\theta}_{2R}' \hat{\theta}_{2R}$ and
\begin{equation*}
  \sigma^2 \equiv \E_R(D_t - \E_R D_t)^2 = 4
  \hat{\theta}_{2R}' \hat{\theta}_{2R} +
  \E_R(\hat{\theta}_{2R}' (I - x_{2t} x_{2t}')
  \hat{\theta}_{2R})^2   \geq 4 \hat{\theta}_{2R}'
  \hat{\theta}_{2R} \quad a.s.
\end{equation*}
Now, $\hat{\theta}_{2R}' \hat{\theta}_{2R}$ is
a.s. positive and uniformly integrable with mean $4 K_2 / (R - K_2 -
1)$ (since $(X_{2R}'X_{2R})^{-1}$ is Inverse-Wishart$(I_{K_2}, R)$),
making $\sigma^2$ uniformly a.s. positive and ensuring that the
distribution of the \oos\ average is not $P^{1/2}$-degenerate.  The
natural \oos\ variance estimator is consistent for $\sigma^2$:
\begin{align*}
  \sh^2 &= P^{-1} \sum_{t=R+1}^{T-1} (D_t - \oosA)^2 \\
  &= 4 \hat{\theta}_{2R}' \Big[P^{-1} \sum_{t=R+1}^{T-1}
  \Big(y_{t+1} x_{2t} - P^{-1} \sum_{s=R+1}^{T-1} y_s x_{2s}\Big)
  \Big(y_{t+1} x_{2t} - P^{-1} \sum_{s=R+1}^{T-1} y_s
  x_{2s}\Big)'\Big] \hat{\theta}_{2R}' \\ & \quad - 4
  \Big[P^{-1} \sum_{t=R+1}^{T-1} \Big(y_{t+1} x_{2t} - P^{-1}
  \sum_{s=R+1}^{T-1} y_s x_{2s}\Big)' \hat{\theta}_{2R}
  \hat{\theta}_{2R}' \Big(x_{2t} x_{2t}' - P^{-1}
  \sum_{s=R+1}^{T-1} x_{2s} x_{2s}' \Big) \hat{\theta}_{2R} \Big]
  \\ &\quad + P^{-1} \sum_{t=R+1}^{T-1}
  \Big[\hat{\theta}_{2R}' \Big(x_{2t} x_{2t}' - P^{-1}
  \sum_{s=R+1}^{T-1} x_{2s} x_{2s}' \Big) \hat{\theta}_{2R}  \Big]^{2}.
\end{align*}
As $P \to \infty$, the first term converges in probability to
$\hat{\theta}_{2R}' \hat{\theta}_{2R}$, the second to zero,
and the third to $\E_R(\hat{\theta}_{2R}'(I - x_{2t}x_{2t}')
\hat{\theta}_{2R})^2$.  Since the number of regressors prevents
$\hat{\theta}_{2R}$ from converging i.p. to zero (its expected
value), $\sh^2$ does not converge to the unconditional
variance.

We can see that the \oos\ average obeys a \clt\ when centered at $\E_R
\bar{D}_R$.  The centered and scaled \oos\ differences satisfy
\begin{equation*}
  \frac{D_t - \E_R D_t}{\sigma} =
  \frac{2 y_{t} x_{2t}' \hat{\theta}_{2R} +
    \hat{\theta}_{2R}' (I - x_t x_t') \hat{\theta}_{2R}}{\sqrt{4
    \hat{\theta}_{2R}' \hat{\theta}_{2R} +
    \E_R(\hat{\theta}_{2R}'(I - x_{2t} x_{2t}') \hat{\theta}_{2R})^2}}.
\end{equation*}
This sequence is not i.i.d. and we can not replace $\hat{\theta}_{2R}$
with a sequence of constants $\theta_{2R}$ since $\hat{\theta}_{2R}$
does not converge to a deterministic limit.  But $\sigma^{-1}
\E_t(\oosA - \E_R\oosA) = 0$, making this sequence an
\mds,\footnote{The sequence is actually exchangeable, but \mds\ is
  enough to establish the \clt.} so it obeys a \clt\ as long as
$\sigma^2$ is uniformly a.s. positive, which we have already
established.  Lemma~\ref{res-mixingale} ensures that the \clt\ holds
more generally.

@@

The first set of simulations study how well the theory in
Section~\ref{sec:oostheory} works in practice.  For each draw of $X$
and $Y$, we construct a one-sided \oos\ interval of the
form
\begin{equation*}
  [ \oosA - 1.28 \hat{\sigma}, \infty) \quad\text{with}\quad
  \sh^2 = \frac1P \sum_{t=R+1}^T (D_t - \oosA)^{2}
\end{equation*}
for $P$ set to every 10th value between 1 and $2T/3$.  We calculate
the percentage of these intervals that contain $\E_R \oosA$ and that
contain $\E_T \oosB$.  Since the data are i.i.d., both of these
quantities are easy to calculate:
\begin{equation*}
  E_R \oosA = \Bigg[\sum_{i=1}^{K_1} (\tilde{\theta}_{iR} - \theta_i)^2 +
  \sum_{i=K_1+1}^{K_2}\theta_i^2\Bigg] - \sum_{i=1}^{K_2} (\hat{\theta}_{iR} -
  \theta_i)^2
\end{equation*}
and
\begin{equation*}
  E_T \oosB = \Bigg[\sum_{i=1}^{K_1} (\tilde{\theta}_{iT} - \theta_i)^2 +
  \sum_{i=K_1+1}^{K_2}\theta_i^2\Bigg] - \sum_{i=1}^{K_2} (\hat{\theta}_{iT} -
  \theta_i)^2
\end{equation*}
where $\tilde{\theta}$ indicates that the coefficient is estimated
using only the regressors in the benchmark model and $\hat{\theta}$
the full model.  We draw 2000 samples for each combination of the
design parameters.

@@

\begin{lem}\label{lem-basic-coupling}
  Suppose Assumptions \ref{asmp-1}--\ref{asmp-3} hold.  Then, for any
  T, $s$, and $t$ with $t \geq s$, there exist random variables
  $D_s^*,\dots,D_t^*$ such that
  \begin{equation}\label{eq:coupling1}
    P[(D_s^*,\dots,D_t^*) \neq (D_s,\dots,D_t)] \leq \beta_{s-u}
  \end{equation}
  and
  \begin{equation}\label{eq:coupling2}
    \E(\phi(D_s^*,\dots, D_t^*) \mid \mathcal{F}_u ) =
    \int
    \phi(D_s,\dots,D_t) f(\mathbf{x}, \mathbf{y})\ d\mathbf{x}\ d\mathbf{y}
  \end{equation}
  almost surely for all measurable functions $\phi$ such that the
  expectations are finite, where
  \[ \mathbf{x} = (x_{s}', \dots, x_{t}')', \qquad
  \mathbf{y} = (y_{s+\h},\dots,y_{t+\h})',\] and $f$ is the
  joint density of $(\mathbf{x}, \mathbf{y})$.  Moreover,
 \begin{equation}\label{eq:coupling3}
   \| D_v^* - D_v \|_2 \leq 2^{1+1/\rho} B_L
   \beta_{s-u}^{\rhoExp}, \qquad v = s,\dots,t.
 \end{equation}
\end{lem}

@@

  Now define
  \begin{equation*}
    D_v^* =
    \begin{cases}
      L(y_{v+\h}^* - x_{1v}^{*\prime} \hat{\theta}_{1R}) - L(y_{v+\h}^* -
      x_{2v}^{*\prime} \hat{\theta}_{2R}) & v \leq T \\
      L(y_{v+\h}^* - x_{1v}^{*\prime} \hat{\theta}_{1T}) - L(y_{v+\h}^* -
      x_{2v}^{*\prime} \hat{\theta}_{2T}) & v > T
    \end{cases}
  \end{equation*}
  with $y_{v+\h}^*$ and $x_{iv}^*$ denoting the elements of $V^*$
  corresponding to $y_{v+\h}$ and $x_{iv}$ in $V_s$.  Equations
  (\ref{eq:coupling1}) and (\ref{eq:coupling2}) are satisfied by
  construction, and (\ref{eq:coupling3}) follows from Lemma
  \ref{lem-extend-mp}.
\end{proof}

>>>>

Both models are estimated using \ols\ with a fixed window:
\[
\hat y_{i,t+\h} = x_{it}'\hat{\theta}_{it}, \qquad i=1,2, \qquad t
= R+1,\dots,T-h
\]
with
\begin{equation*}
  \bh{it} =
  \begin{cases}
    \big(\sum_{s=1}^{R-\h} x_{is}x_{is}'\big)^{-1} \sum_{s=1}^{R-\h}
    x_{is} y_{s+\h} & t \leq T \\
    \big(\sum_{s=1}^{T-\h} x_{is}x_{is}'\big)^{-1}
    \sum_{s=1}^{T-\h} x_{is} y_{s+\h} & t > T.
  \end{cases}
\end{equation*}
The observed \oos\ average is denoted $\oosA$: \[\oosA \equiv P^{-1}
\oosSum{t}{1} D_{Rt}\]  with
\[
D_{Rt} \equiv L(y_{t+\h} - x_{1t}'\bh{1t}) - L(y_{t+\h} - x_{2t}'\bh{2t}).
\]
The \oos\ statistic of interest is $\oost \equiv \sqrt{P} \oosA /
\hat\sigma_R$, for $\hat\sigma_R^2$ some (possibly Heteroskedasticity- and
Autocorrelation-Consistent, or \hac) estimator of the asymptotic
variance of $\oosA$.

As discussed above, the forecaster's goal is to choose the model with
lower average future loss; i.e. to determine whether $\oosB$ is
positive or negative,with
\[
\oosB \equiv Q^{-1} \sum_{t=T+1}^{T+Q} D_{Tt}, \quad
D_{Tt} \equiv L(y_{t+h} - x_{1t}' \bh{1t}) - L(y_{t+h} - x_{2t}' \bh{2t}).
\]
Note that the coefficients are reestimated with the full dataset.
Since the forecaster's information set, $\mathcal{F}_T$, does not
contain $\oosB$, he or she should decide between the models based on
the conditional expectation, $\E_T \oosB$. If this quantity is
positive, the first model is expected to perform the worst in the
future, and if it is negative the second is.

@@

\begin{asmp}\label{asmp-3}
  The loss function $L$ is continuous and has finite left and right
  derivatives almost everywhere.  Also, $L(0) = 0$, and there is a
  constant $B_L$ such that $\|D_{R,R+j}\|_\rho \leq B_L$ and
  $\|D_{T,T+j}\|_\rho \leq B_L$ for all $j \geq 1$ ($\rho$ is defined
  in Assumption~\ref{asmp-1}); $\| x_{it}' \hat{\theta}_{is}^R \|_2 \leq
  B_L$ and $\| x_{it}' \hat{\theta}_{is}^T \|_2 \leq B_L$; and
  \begin{equation*}
    \| L'(y_{t+h} - x_{it}'\tilde{\theta}_{it}) \|_2 \leq B_L
  \end{equation*}
  for $\tilde{\theta}_{it}$ a weighted average of any two estimates
  $\hat{\theta}_{is}^R$ and $\hat{\theta}_{is'}^R$ or
  $\hat{\theta}_{is}^T$ and $\hat{\theta}_{is'}^T$ with $s, s' \leq t$.
\end{asmp}

@@

\begin{thm}\label{res-confidence-intervals}
  Suppose that Assumptions \ref{asmp-1}--\ref{asmp-4} hold, that
  $\hat\sigma_R^2$ is an estimator satisfying
  \[
    \hat\sigma_R^2 - \sigma_R^2 \to^p 0 \quad \text{with}\quad
    \sigma_R^2 \equiv \var_R(\sqrt{P} \oosA) \equiv P \E_R(\bar{D}_R -
    \E_R \bar{D}_R)^2,
  \]
  and that $\sigma_R^2$ is uniformly a.s. positive.  Then
  \[
  \frac{\oosA - \E_T\oosB}{\hat\sigma_R / \sqrt{P}}
  \to^d N(0,1).
  \]
  Consequently, each of the usual Gaussian confidence intervals,
  \begin{gather}
  [\oosA - z_{\alpha/2} \, \hat\sigma_R /
      \sqrt{P}, \oosA + z_{\alpha/2} \hat\sigma_R / \sqrt{P}],\label{interval-twosided} \\
  [\oosA - z_{\alpha} \, \hat\sigma_R / \sqrt{P}, +\infty), \label{interval-greater}
  \intertext{and}
(-\infty, \oosA + z_{\alpha}
      \, \hat\sigma_R / \sqrt{P}],
    \end{gather}
    contains $\E_T\oosB$ with probability $\alpha$ in the limit,
  with $z_{\alpha}$ the $1-\alpha$ quantile of the standard normal
  distribution.
\end{thm}
The requirement that $\sigma_R^2$ be uniformly a.s. positive is not
restrictive under our asymptotic theory.  Since the models'
coefficients are estimated with uncertainty in the limit, the two
models give different forecasts even if they both nest the \dgp.  This
is intuitively similar to \citepos{GiW:06} rolling-window result, but
comes from different asymptotic theory; Giacomini and White keep the
variance of the \oos\ average positive by letting $P \to \infty$ with
$R$ fixed; in this paper, $R \to \infty$ but the variance remains
positive since $K \to \infty$ (quickly) as well.

This result allows forecasters to use the \dmw\ test to conduct
inference about the models' accuracy.  Since the \oos\ average,
$\oosA$, is asymptotically normal and centered on the expected future
performance, $t$-tests constructed from the \oos\ errors necessarily
test hypotheses about $\E_T \oosB$ and not the performance of the
pseudo-true models or the unconditional performance of the models.
This application to testing can be expressed as its own theorem:
\begin{thm}\label{res:oostest}
Suppose that the conditions of Theorem \ref{res-confidence-intervals}
hold and that $\lim \Pr[\E_T \oosB \leq 0] > 0$.  Then
\begin{equation}
  \lim \Pr[P^{1/2}\oosA/\hat\sigma_R > z_{\alpha} \mid \E_T
  \oosB \leq 0] \leq \alpha.
\end{equation}
\end{thm}

Two intermediate Lemmas will help us understand Theorem
\ref{res-confidence-intervals} better.  Its proof consists of two
steps: we can write
\begin{equation}
  \sqrt{P} (\oosA - \E_T \oosB) = \sqrt{P} (\oosA - \E_R
  \oosA) + \sqrt{P} (\E_R \oosA - \E_T \oosB),
\end{equation}
so Theorem~\ref{res-confidence-intervals} holds if
\begin{equation}\label{coefpart}
  \E_R \oosA = \E_T \oosB + o_p(P^{-1/2}).
\end{equation}
and
\begin{equation}\label{normalpart}
  \sqrt{P}\frac{\oosA - \E_R \oosA}{\sigma_R} \to^d N(0,1)
\end{equation}
The first Lemma establishes~\eqref{coefpart} and the second shows that
the process $\{D_{Rt} - \E_R \oosA\}$ is an $L_2$-mixingale of size
$-1/2$ that satisfies a \clt, ensuring~\eqref{normalpart}.

@@

Note that $\E_T \oosB - \E_R \oosA$ is noise introduced by
approximating the performance of the full-sample estimates with that
of the partial-sample estimates.  Moreover, the $O_p(\sqrt{P/T})$ term
is binding; it is straightforward to find examples for which
$\sqrt{T/P}(\E_T \oosB - \E_R \oosA)$ does not converge to
zero.\footnote{Notice that, if the observations are i.i.d. and
  $(y_{t+h}, x_{it}) \sim N(0,I)$, $(\hat{\theta}_{iR}^R -
  \hat{\theta}_{iT}^T) \sim N(0, (X_{iR}'X_{iR})^{-1} -
  (X_{iT}'X_{iT})^{-1})$ given $X_{iT}$, and so
  $\var(x_{i,T+1}'\hat{\theta}_{iR}^R - x_{i,T+1}'
  \hat{\theta}_{iT}^T) \gg P K_i / (R - K_2)(T - K_2)$.}
Consequently, there is little scope to relax our assumption that
$P^2/T \to 0$.  If $\lim P^2/T$ is positive and finite then there is
the possibility of adjusting the asymptotic distribution of the test
statistic to account for this noise term as in \citet{Wes:96}, but we
leave that to future research.  If $P^2/T \to \infty$, which is
closest to common practice in applied research, the noise term
dominates the asymptotic distribution of the \oos\ $t$-test and the
test has only trivial power in the limit if it has correct size.

\begin{lem}\label{res-mixingale}
  Part 1: If Assumptions~\ref{asmp-1}--\ref{asmp-3} hold then the
  inequality
  \begin{equation}\label{mixingaleR}
    \lVert \E_{R+j-l}(D_{R,R+j} - \E_R D_{R,R+j}) \rVert_2 \leq
    \couplingConstant \;
    \zeta_l
  \end{equation}
  with $\zeta_l = O(l^{-1/2 - \delta})$ holds for some positive
  $\delta$, any positive $j$, and any $l$ between 0 and $j$.  As a
  consequence, $\{D_{Rt} - \E_R D_{Rt}, \mathcal F_t\}$ is an
  $L_2$-mixingale of size $-1/2$.

  \noindent Part 2: Moreover,~\eqref{normalpart} holds under
  Assumptions~\ref{asmp-1}--\ref{asmp-4} if $\sigma_R^2$ is
  a.s. uniformly positive.
\end{lem}

@@

Finally, for Theorem \ref{res-confidence-intervals} to be useful, we
need a consistent estimator of $\sigma_R^2$. Lemma
\ref{res-variance-estimator} establishes that the usual \oos\ \hac\
variance estimators are consistent.\footnote{As with
      Lemma~\ref{res-mixingale}, the proof of this lemma takes
      \citepos{JoD:00} existing consistency results
      for \ned\ processes and shows that the same arguments can be
      applied to our \oos\ mixingale.}

\begin{lem}
  \label{res-variance-estimator} Suppose Assumptions
  \ref{asmp-1}--\ref{asmp-5} hold and that $\gamma \to \infty$ and
  $\gamma/P \to 0$ as $T \to \infty$, and define
  \begin{equation*}
    \hat{\sigma}_R^2 \equiv P^{-1} \oosSum{s,t}{1} (D_{Rt} - \oosA)(D_{Rs} - \oosA)
    \vWeight.
  \end{equation*}
  Then $\hat{\sigma}_R^2 - \sigma_R^2 \to^p 0$.
\end{lem}

@@

The first two conditional moments of $D_t$ for $t > R$ (given
$\mathcal{F}_R$) are straightforward to derive: $\E_R D_{Rt} = -
\hat{\theta}_{2R}' \hat{\theta}_{2R}$ and
\begin{equation*}
  \sigma_R^2 \equiv \E_R(D_{Rt} - \E_R D_{Rt})^2 = 4
  \hat{\theta}_{2R}^{R\prime} \hat{\theta}_{2R}^{R} +
  \E_R(\hat{\theta}_{2R}^{R\prime} (I - x_{2t} x_{2t}')
  \hat{\theta}_{2R}^{R})^2   \geq 4 \hat{\theta}_{2R}^{R\prime}
  \hat{\theta}_{2R}^R \quad a.s.
\end{equation*}
Now, $\hat{\theta}_{2R}' \hat{\theta}_{2R}$ is
a.s. positive and uniformly integrable with mean $4 K_2 / (R - K_2 -
1)$ (since $(X_{2R}'X_{2R})^{-1}$ is Inverse-Wishart$(I_{K_2}, R)$),
making $\sigma_R^2$ uniformly a.s. positive and ensuring that the
distribution of the \oos\ average is not $P^{1/2}$-degenerate.  The
natural \oos\ variance estimator is consistent for $\sigma_R^2$:
\begin{align*}
  \hat{\sigma}_R^2 &= P^{-1} \sum_{t=R+1}^{T-1} (D_{Rt} - \oosA)^2 \\
  &= 4 \hat{\theta}_{2R}^{R\prime} \Big[P^{-1} \sum_{t=R+1}^{T-1}
  \Big(y_{t+1} x_{2t} - P^{-1} \sum_{s=R+1}^{T-1} y_s x_{2s}\Big)
  \Big(y_{t+1} x_{2t} - P^{-1} \sum_{s=R+1}^{T-1} y_s
  x_{2s}\Big)'\Big] \hat{\theta}_{2R}' \\ & \quad - 4
  \Big[P^{-1} \sum_{t=R+1}^{T-1} \Big(y_{t+1} x_{2t} - P^{-1}
  \sum_{s=R+1}^{T-1} y_s x_{2s}\Big)' \hat{\theta}_{2R}
  \hat{\theta}_{2R}' \Big(x_{2t} x_{2t}' - P^{-1}
  \sum_{s=R+1}^{T-1} x_{2s} x_{2s}' \Big) \hat{\theta}_{2R} \Big]
  \\ &\quad + P^{-1} \sum_{t=R+1}^{T-1}
  \Big[\hat{\theta}_{2R}' \Big(x_{2t} x_{2t}' - P^{-1}
  \sum_{s=R+1}^{T-1} x_{2s} x_{2s}' \Big) \hat{\theta}_{2R}'  \Big]^{2}.
\end{align*}
As $P \to \infty$, the first term converges in probability to
$\hat{\theta}_{2R}^{R\prime} \hat{\theta}_{2R}^R$, the second to zero,
and the third to $\E_R(\hat{\theta}_{2R}^{R\prime}(I - x_{2t}x_{2t}')
\hat{\theta}_{2R}^R)^2$.  Since the number of regressors prevents
$\hat{\theta}_{2R}^R$ from converging i.p. to zero (its expected
value), $\hat{\sigma}_R^2$ does not converge to the unconditional
variance.

We can see that the \oos\ average obeys a \clt\ when centered at $\E_R
\bar{D}_R$.  The centered and scaled \oos\ differences satisfy
\begin{equation*}
  \frac{D_{Rt} - \E_R D_{Rt}}{\sigma_R} =
  \frac{2 y_{t} x_{2t}' \hat{\theta}_{2R} +
    \hat{\theta}_{2R}' (I - x_t x_t') \hat{\theta}_{2R}}{\sqrt{4
    \hat{\theta}_{2R}' \hat{\theta}_{2R} +
    \E_R(\hat{\theta}_{2R}'(I - x_{2t} x_{2t}') \hat{\theta}_{2R})^2}}.
\end{equation*}
This sequence is not i.i.d. and we can not replace $\hat{\theta}_{2R}$
with a sequence of constants $\theta_{2R}$ since $\hat{\theta}_{2R}$
does not converge to a deterministic limit.  But $\sigma_R^{-1}
\E_t(\oosA - \E_R\oosA) = 0$, making this sequence an
\mds,\footnote{The sequence is actually exchangeable, but \mds\ is
  enough to establish the \clt.} so it obeys a \clt\ as long as
$\sigma_R^2$ is uniformly a.s. positive, which we have already
established.  Lemma~\ref{res-mixingale} ensures that the \clt\ holds
more generally.

@@

The first set of simulations study how well the theory in
Section~\ref{sec:oostheory} works in practice.  For each draw of $X$
and $Y$, we construct a one-sided \oos\ interval of the
form
\begin{equation*}
  [ \oosA - 1.28 \hat{\sigma_R}, \infty) \quad\text{with}\quad
  \hat\sigma_R^2 = \frac1P \sum_{t=R+1}^T (D_t - \oosA)^{2}
\end{equation*}
for $P$ set to every 10th value between 1 and $2T/3$.  We calculate
the percentage of these intervals that contain $\E_R \oosA$ and that
contain $\E_T \oosB$.  Since the data are i.i.d., both of these
quantities are easy to calculate:
\begin{equation*}
  E_R \oosA = \Bigg[\sum_{i=1}^{K_1} (\tilde{\theta}_{iR} - \theta_i)^2 +
  \sum_{i=K_1+1}^{K_2}\theta_i^2\Bigg] - \sum_{i=1}^{K_2} (\hat{\theta}_{iR} -
  \theta_i)^2
\end{equation*}
and
\begin{equation*}
  E_T \oosB = \Bigg[\sum_{i=1}^{K_1} (\tilde{\theta}_{iT} - \theta_i)^2 +
  \sum_{i=K_1+1}^{K_2}\theta_i^2\Bigg] - \sum_{i=1}^{K_2} (\hat{\theta}_{iT} -
  \theta_i)^2
\end{equation*}
where $\tilde{\theta}$ indicates that the coefficient is estimated
using only the regressors in the benchmark model and $\hat{\theta}$
the full model.  We draw 2000 samples for each combination of the
design parameters.

@@

\begin{lem}\label{lem-basic-coupling}
  Suppose Assumptions \ref{asmp-1}--\ref{asmp-3} hold.  Then, for any
  T, $s$, and $t$ with $t \geq s$, there exist random variables
  $\tilde D_s,\dots,\tilde D_t$ such that
  \begin{equation}\label{eq:coupling1}
    P[(\tilde D_s,\dots,\tilde D_t) \neq (D_s,\dots,D_t)] \leq \beta_{s-u}
  \end{equation}
  and
  \begin{equation}\label{eq:coupling2}
    \E(\phi(\tilde D_s,\dots, \tilde D_t) \mid \mathcal{F}_u ) =
    \int
    \phi(D_s,\dots,D_t) f(\mathbf{x}, \mathbf{y})\ d\mathbf{x}\ d\mathbf{y}
  \end{equation}
  almost surely for all measurable functions $\phi$ such that the
  expectations are finite, where
  \[ \mathbf{x} = (x_{s}', \dots, x_{t}')', \qquad
  \mathbf{y} = (y_{s+\h},\dots,y_{t+\h})',\] and $f$ is the
  joint density of $(\mathbf{x}, \mathbf{y})$.  Moreover,
 \begin{equation}\label{eq:coupling3}
   \| \tilde D_v - D_v \|_2 \leq 2^{1+1/\rho} B_L
   \beta_{s-u}^{\rhoExp}, \qquad v = s,\dots,t.
 \end{equation}
\end{lem}

@@

  Now define
  \begin{equation*}
    \tilde D_v =
    \begin{cases}
      L(y_{v+\h}^* - x_{1v}^{*\prime} \hat{\theta}_{1R}) - L(y_{v+\h}^* -
      x_{2v}^{*\prime} \hat{\theta}_{2R}) & v \leq T \\
      L(y_{v+\h}^* - x_{1v}^{*\prime} \hat{\theta}_{1T}) - L(y_{v+\h}^* -
      x_{2v}^{*\prime} \hat{\theta}_{2T}) & v > T
    \end{cases}
  \end{equation*}
  with $y_{v+\h}^*$ and $x_{iv}^*$ denoting the elements of $V^*$
  corresponding to $y_{v+\h}$ and $x_{iv}$ in $V_s$.  Equations
  (\ref{eq:coupling1}) and (\ref{eq:coupling2}) are satisfied by
  construction, and (\ref{eq:coupling3}) follows from Lemma
  \ref{lem-extend-mp}.
\end{proof}
